---
title: "Biostat HW2 Prob 6"
author: "Alyssa Vanderbeek"
date: "9/25/2018"
output: pdf_document
---

## Problem 6
A researcher is conducting a study to examine associations of depression and cognitive performance with migraine symptoms. Use data from the Human Epilepsy Project (HEP) to answer the following questions.

Migraine status 0-no, 1-yes
NDDIE: Neurological Disorder Depression Inventory for Epilepsy
CESD: Center for Epidemiologic Studies Depression Scale
Cognitive evaluation: Aldenkamp-Baker

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

migraine_data = 
  readxl::read_excel('~/Migraine.xlsx') %>%
  janitor::clean_names() %>%  # brings colnames to lowercase, no white space
  mutate(migraine = recode(migraine, `0` = 'no', `1` = 'yes'))
head(migraine_data)
```


(a) Summarize the three variables above for epilepsy patients with and without migraine. Carefully choose the descriptive statistics and report both measures of location/spread, sample sizes (N) and number of missing values for each variable.
For NDDIE and CSED, use the original scores and the following cutoffs: NDDIE (cutoff of 16), CSED (cutoff of 16). 


## CESD summary
```{r}
migraine_data %>%
  group_by(migraine) %>%
  summarise(n_cesd = table(migraine),  
            avg_cesd = mean(cesd, na.rm = T),  
            quartile_25 = quantile(cesd, na.rm = T)[2],
            median = median(cesd, na.rm = T),
            quartile_75 = quantile(cesd, na.rm = T)[4],
            sd = sd(cesd, na.rm = T))
```

## NDDIE summary
```{r}
migraine_data %>%
  group_by(migraine) %>%
  summarise(n_nddie = table(migraine),  # number of subjects per group
            avg_cesd = mean(nddie, na.rm = T),  # mean CESD
            quartile_25 = quantile(nddie, na.rm = T)[2],
            median = median(nddie, na.rm = T),
            quartile_75 = quantile(nddie, na.rm = T)[4],
            sd = sd(nddie, na.rm = T))
```


```{r}
migraine_recode = 
  migraine_data %>% 
  mutate(cesd_geq16 = ifelse(cesd >= 16, 'yes', 'no'),
         nddie_geq16 = ifelse(nddie >= 16, 'yes', 'no'))
```

Create graphical displays to show the scores distributions for NDDIE, CSED, ABNAS (memory and language) by group (migraine vs no-migraine). Please add your recommendations / comments. (5p)

